---
title: "STAT 426 Project 1"
author: "Corey Huang"
date: '2022-04-11'
output:
  pdf_document: default
  word_document: default
---

## Import the Data  
```{r}
library(faraway)
library(VGAM)
library(tidyverse)
keto <- read.csv("ketorolac.csv")
```

## Data Visualization  
```{r}
plot(keto$LongStay~TOR, data = keto)
plot(keto$BadPain~TOR, data=keto)
hist(keto$LongStay, data=keto)
hist(keto$BadPain, data=keto)
```


## Outliers  
```{r}
outliers <- boxplot(keto$Morphine, plot = FALSE)$out
outliers
v = c(which(keto$Morphine == 317.5), which(keto$Morphine == 310.3), which(keto$Morphine == 300.4))
keto1 <- keto[-v,]
keto1 = na.omit(keto1)
```


## Model Fitting  
```{r}
full_model <- glm(cbind(LongStay, BadPain) ~., family = binomial, data = keto)
summary(full_model)
```

## Odds-Ratio for each predictor  
```{r}
log_odds <- full_model$coefficients
exp(full_model$coefficients)
log_conf <- confint(full_model)
exp(confint(full_model))
```


# Variable Selection  
```{r}
step(full_model, direction = "backward", trace = FALSE)
step(full_model, direction = "forward", trace = FALSE)
step(full_model, direction = "both", trace = FALSE)
```


## Confounding Variables
```{r}
model.yr <- glm(TOR~DOS.yr, family="binomial", data=keto)
exp(cbind(coef(model.yr), confint(model.yr, level = 0.95)))
model.stage <- glm(TOR~Stage, family="binomial", data=keto)
exp(cbind(coef(model.stage), confint(model.stage, level = 0.95)))
model.diagnosis <- glm(TOR~Diagnosis, family="binomial", data=keto)
exp(cbind(coef(model.diagnosis), confint(model.diagnosis, level = 0.95)))
model.duration <- glm(TOR~duration, family="binomial", data=keto)
exp(cbind(coef(model.duration), confint(model.duration, level = 0.95)))
model.age <- glm(TOR~Age, family="binomial", data=keto)
exp(cbind(coef(model.age), confint(model.age, level = 0.95)))
model.addpro <- glm(TOR~AddPro, family="binomial", data=keto)
exp(cbind(coef(model.addpro), confint(model.addpro, level = 0.95)))
model.HYS <- glm(TOR~HYS, family="binomial", data=keto)
exp(cbind(coef(model.HYS), confint(model.HYS, level = 0.95)))
model.OV <- glm(TOR~OV, family="binomial", data=keto)
exp(cbind(coef(model.OV), confint(model.OV, level = 0.95)))
model.LNS <- glm(TOR~LNS, family="binomial", data=keto)
exp(cbind(coef(model.LNS), confint(model.LNS, level = 0.95)))
model.Morphine <- glm(TOR~Morphine, family="binomial", data=keto)
exp(cbind(coef(model.Morphine), confint(model.Morphine, level = 0.95)))
model.Comps <- glm(TOR~Comps, family="binomial", data=keto)
exp(cbind(coef(model.Comps), confint(model.Comps, level = 0.95)))
```


# Re-Fit the Model
```{r}
reduced_model_logit <- glm(cbind(LongStay, BadPain) ~ General.diet + TOR + Comps + Morphine, family = binomial(link = ("logit")), data = keto1)
reduced_model_probit <- glm(cbind(LongStay, BadPain) ~ General.diet + TOR + Comps + Morphine, family = binomial(link = ("probit")), data = keto1)
reduced_model_cloglog <- glm(cbind(LongStay, BadPain) ~ General.diet + TOR + Comps + Morphine, family = binomial(link = ("cloglog")), data = keto1)
```

# Diagnostics  
## Cook's Distance  
```{r}
cooks_red <- cooks.distance(reduced_model_cloglog)
cooks_dist <- cooks_red > (4 / length(cooks_red))
cooks_red[cooks_dist]
x=c(31, 56)
keto2 <- keto1[-x,]
keto2 <- na.omit(keto2)
```

## Highly Influential Points
```{r}
df_red = dfbeta(reduced_model_cloglog)
abs(df_red) > 1
```

# ROC Curves
## ROC Curve - LongStay on Logit Model
```{r}
pihat <- fitted(reduced_model_logit)
false.neg <- c(0,cumsum(tapply(keto1$LongStay,pihat,sum)))
true.neg <- c(0,cumsum(table(pihat))) - false.neg
plot(1-true.neg/max(true.neg), 1-false.neg/max(false.neg), type="l",
main="ROC Curve", xlab="1 - Specificity", ylab="Sensitivity",
xlim=c(0,1), ylim=c(0,1), lwd=3)
abline(a=0, b=1, lty=2, col="blue")
mean(outer(pihat[keto1$LongStay==1], pihat[keto1$LongStay==0], ">")
+ 0.5 * outer(pihat[keto1$LongStay==1], pihat[keto1$LongStay==0], "=="))
```
Area under ROC Curve: 0.8302969

## ROC Curve - BadPain on Logit Model
```{r}
pihat <- fitted(reduced_model_logit)
false.neg <- c(0,cumsum(tapply(keto1$BadPain,pihat,sum)))
true.neg <- c(0,cumsum(table(pihat))) - false.neg
plot(1-true.neg/max(true.neg), 1-false.neg/max(false.neg), type="l",
main="ROC Curve", xlab="1 - Specificity", ylab="Sensitivity",
xlim=c(0,1), ylim=c(0,1), lwd=3)
abline(a=0, b=1, lty=2, col="blue")
mean(outer(pihat[keto1$BadPain==1], pihat[keto1$BadPain==0], ">")
+ 0.5 * outer(pihat[keto1$BadPain==1], pihat[keto1$BadPain==0], "=="))
```
Area under ROC Curve: 0.3941468

## ROC Curve - LongStay on Probit Model
```{r}
pihat <- fitted(reduced_model_probit)
false.neg <- c(0,cumsum(tapply(keto1$LongStay,pihat,sum)))
true.neg <- c(0,cumsum(table(pihat))) - false.neg
plot(1-true.neg/max(true.neg), 1-false.neg/max(false.neg), type="l",
main="ROC Curve", xlab="1 - Specificity", ylab="Sensitivity",
xlim=c(0,1), ylim=c(0,1), lwd=3)
abline(a=0, b=1, lty=2, col="blue")
mean(outer(pihat[keto1$LongStay==1], pihat[keto1$LongStay==0], ">")
+ 0.5 * outer(pihat[keto1$LongStay==1], pihat[keto1$LongStay==0], "=="))
```
Area under ROC Curve: 0.830072

## ROC Curve - BadPain on Probit Model
```{r}
pihat <- fitted(reduced_model_probit)
false.neg <- c(0,cumsum(tapply(keto1$BadPain,pihat,sum)))
true.neg <- c(0,cumsum(table(pihat))) - false.neg
plot(1-true.neg/max(true.neg), 1-false.neg/max(false.neg), type="l",
main="ROC Curve", xlab="1 - Specificity", ylab="Sensitivity",
xlim=c(0,1), ylim=c(0,1), lwd=3)
abline(a=0, b=1, lty=2, col="blue")
mean(outer(pihat[keto1$BadPain==1], pihat[keto1$BadPain==0], ">")
+ 0.5 * outer(pihat[keto1$BadPain==1], pihat[keto1$BadPain==0], "=="))
```
Area under ROC Curve: 0.3953116

## ROC Curve - LongStay on Cloglog Model
```{r}
pihat <- fitted(reduced_model_cloglog)
false.neg <- c(0,cumsum(tapply(keto1$LongStay,pihat,sum)))
true.neg <- c(0,cumsum(table(pihat))) - false.neg
plot(1-true.neg/max(true.neg), 1-false.neg/max(false.neg), type="l",
main="ROC Curve", xlab="1 - Specificity", ylab="Sensitivity",
xlim=c(0,1), ylim=c(0,1), lwd=3)
abline(a=0, b=1, lty=2, col="blue")
mean(outer(pihat[keto1$LongStay==1], pihat[keto1$LongStay==0], ">")
+ 0.5 * outer(pihat[keto1$LongStay==1], pihat[keto1$LongStay==0], "=="))
```
Area under ROC Curve: 0.8359199

## ROC Curve - BadPain on Cloglog Model
```{r}
pihat <- fitted(reduced_model_cloglog)
false.neg <- c(0,cumsum(tapply(keto1$BadPain,pihat,sum)))
true.neg <- c(0,cumsum(table(pihat))) - false.neg
plot(1-true.neg/max(true.neg), 1-false.neg/max(false.neg), type="l",
main="ROC Curve", xlab="1 - Specificity", ylab="Sensitivity",
xlim=c(0,1), ylim=c(0,1), lwd=3)
abline(a=0, b=1, lty=2, col="blue")
mean(outer(pihat[keto1$BadPain==1], pihat[keto1$BadPain==0], ">")
+ 0.5 * outer(pihat[keto1$BadPain==1], pihat[keto1$BadPain==0], "=="))
```
Area under ROC Curve: 0.3976412

Average area under ROC for each link function:
Logit: 0.61222185
Probit: 0.6126918
Cloglog: 0.61678055

# Fit the Final Model
```{r}
reduced_model_cloglog <- glm(cbind(LongStay, BadPain) ~ General.diet + TOR + Comps + Morphine, family = binomial(link = ("cloglog")), data = keto2)
summary(reduced_model_cloglog)
```


## Odds Ratios of New Reduced Model
```{r}
log_odds <- reduced_model_cloglog$coefficients
exp(log_odds)
log_conf <- confint(reduced_model_cloglog)
exp(confint(reduced_model_cloglog))
```
Since the odds-ratio for Toradol is 0.6607466, we can say that Toradol decreases LongStay and BadPain by a factor of 0.6607466.

# Overdispersion

```{r}
deviance(reduced_model_cloglog)/df.residual(reduced_model_cloglog) 
```
Since the ratio is not larger than 1, we can say that overdispersion does not occur in the model, however underdispersion may possibly be an issue. 


